Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2012-03-08 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2012:03:08	13:18:56	(Thu 8.Mar)
la2pdf commit - 2012:03:08	14:02:17	(Thu 8.Mar)
la2pdf commit - 2012:03:08	14:33:11	(Thu 8.Mar)

@@

\begin{itemize}
\item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
\item But: Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust. Computationally complex O($M^3$), and although subspace/beamspace methods exist, 
\item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
\end{itemize}

The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.

While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output.

We have implemented a Low Complexity Adaptive (LCA) beamformer, which adaptively selects a window from a predefined set. The set is comprised of windows that are typical solutions found by the Minimum Variance method. The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.

An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.
\end{abstract}

@@

\begin{itemize}
\item Sell it!
\item As briefly as possible, introduce beamforming.
\item Adaptive beamformer potential
\item Why active sonar make it hard. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
\item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was that guy in the 80ties?}
\item Outline: Start with Capon's response when means for robustification are added.
\end{itemize}

\ \\
\IEEEPARstart{T}{o} form images from a modern phased array sonar system the received wavefield is usually recorded, and then postprocessed by a digital beamformer. The beamformer applies delays and weights to the sensor channels, the beamformer adjusts the arrays spatial response to focus at one pixel at a time.  such that signals emanating from regions of interest add constructively, while ensuring that noise and interference from other angles do not.

The imaging capabilities of a modern phased array sonar system depend on physical attributes such element response and array geometry, the transmitted signal, as well as the beamforming method being used on transmission and reception. Beamforming is the concept of applying delays and weights to the sensors channels to steer the arrays response to points of interest.
% \item Why beamforming?
% \item Adaptive beamformers
% \item Challenges in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
% \item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was first?}
% \item Outline: Start with Capon's response when means for robustification are added.

>>>>

\usepackage{tikz}
\usetikzlibrary{shapes,snakes}
\usepackage{amsmath,amssymb}

\newenvironment{outline}
{\begin{itemize}}
{\end{itemize}}

%    \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
%    \setlength\marginparsep{3pt}
%    \setlength\marginparwidth{42pt}
%    \newlength\marginparwidthsmall
%    \setlength\marginparwidthsmall{\marginparwidth}
%    \addtolength\marginparwidthsmall{-7pt}
%    \newcommand\todo[1]{%
%       \addtocounter{todoidx}{1}%
%       {\color{Red}\fbox{\bf\thetodoidx{}}}%
%       \marginpar{%
%          {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
%          \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{#1}}}}
%

@@

\begin{itemize}
\item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
\item But:
\begin{itemize}
\item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
\item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
\item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
\end{itemize}
\item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
\end{itemize}

\end{abstract}

@@

\begin{itemize}
\item Sell it!
\item As briefly as possible, introduce beamforming.
\item Adaptive beamformer potential
\item Why adaptive beamformers struggle in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive. Constraints must be applied in one way or another - parameters must be tuned.
\item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was that guy in the 80ties?}
\item Outline: 
\begin{itemize}
\item Methods.
\item See what happens to Capon's response when means for robustification are added.
\item Setup the window set for LCA to match this.
\item Notice that the LCA performs like Capon, even with a small window database.
\item Start with Kaiser, see that other window functions also work well.
\end{itemize}
\end{itemize}

\section{Methods}

\begin{align}
z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\X_\Delta[n]
\end{align}
where
\begin{align}
\w[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]} \qquad \text{and} \X_\Delta[n] = 
\end{align}

\section{Conclusion}

\begin{itemize}
\item Something similar to the abstract.
\end{itemize}

% \ \\
% \IEEEPARstart{T}{o} form images from a modern phased array sonar system the received wavefield is usually recorded, and then postprocessed by a digital beamformer. The beamformer applies delays and weights to the sensor channels, the beamformer adjusts the arrays spatial response to focus at one pixel at a time.  such that signals emanating from regions of interest add constructively, while ensuring that noise and interference from other angles do not. 
% 
% The imaging capabilities of a modern phased array sonar system depend on physical attributes such element response and array geometry, the transmitted signal, as well as the beamforming method being used on transmission and reception. Beamforming is the concept of applying delays and weights to the sensors channels to steer the arrays response to points of interest.
